---
title: "getting-started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# gimap tutorial

```{r}
library(magrittr)
library(gimap)
```

```{r}
example_data <- example_data()

# Let's examine this example metadata
example_data
```

```{r}
names(example_data)
```

`example_data` is a named list storing count data and ID mapping data....

```{r}
colnames(example_data$countdata)
```

This count data has ...

```{r}
colnames(example_data$idmapping)
```

And the ID mapping data has ...

## Setting up data

We're going to set up three datasets. Two are required; the first is the counts that the genetic interaction analysis will be used for. 

```{r}
example_counts <- example_data$countdata %>%
  dplyr::select(c("Day00_RepA", "Day05_RepA", "Day22_RepA", "Day22_RepB", "Day22_RepC")) %>%
  as.matrix()
```

The next datasets are metadata that describe the dimensions of the count data. 
- These both need to be data frames. 
- The sizes of these metadata must correspond to the dimensions of the counts data. 
- The first column of the pg_metadata must be a unique id

One of these (`example_pg_metadata`) is required because it is necessary to know the IDs and be able to map them to pgRNA constructs. 

```{r}
# pg metadata is the information that describes the paired guide RNA targets
example_pg_metadata <- example_data$countdata %>%
  dplyr::select("id")

# sample metadata is the information that describes 
example_sample_metadata <- data.frame(
  id = 1:5,
  day = as.factor(c("Day00", "Day05", "Day22", "Day22", "Day22")), 
  rep = as.factor(c("RepA", "RepA", "RepA", "RepB", "RepC"))
)
```

We'll need to provide `example_counts`, `pg_metadata` and `example_data$idmapping` to `setup_data()`. We can provide `sample_metadata`, but it is not required. (but should it be???)

Now let's setup our data using `setup_data()`. Optionally we can provide the metadata in this function as well so that it is stored with the data. 

```{r}
gimap_dataset <- setup_data(counts = example_counts, 
                            pg_metadata = example_pg_metadata,
                            sample_metadata = example_sample_metadata)
```

You'll notice that this set up gives us a list of formatted data. This contains the original counts we gave `setup_data()` function but also normalized counts, and the total counts per sample. 

```{r}
str(gimap_dataset)
```

```{r}
sessionInfo()
```
