---
title: "getting-started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# gimap tutorial

```{r}
library(magrittr)
library(gimap)
```

```{r}
example_data <- example_data()

# Let's examine this example metadata

# TODO: ## this may or may not be the right place to do it, but perhaps a more descriptive comment of what we're looking at would help. e.g. "Let's examine this example pgPEN counts table, with column IDs representing time points at which gDNA was extracted and the experimental (biological/technical?) replicate number" 

example_data
```

```{r}
colnames(example_data)
```

## Setting up data

We're going to set up three datasets. The first is required, it's the counts that the genetic interaction analysis will be used for. 

```{r}
# TODO: ## perhaps here we can clarify the flexibility of the workflow / what can be set up. e.g. does this function allow exploration / comparison of specific days and reps? or is that not really needed for the purposes of the vignette?

example_counts <- example_data %>%
  dplyr::select(c("Day00_RepA", "Day05_RepA", "Day22_RepA", "Day22_RepB", "Day22_RepC")) %>%
  as.matrix()
```

The next two datasets are metadata that describe the dimensions of the count data. 
- These both need to be data frames. 
- The sizes of these metadata must correspond to the dimensions of the counts data. 
- The first column of the pg_metadata must be a unique id

```{r}
# pg metadata is the information that describes the paired guide RNA targets

# TODO: ## this is really minor but I think a more accurate description would be closer to: "pg metadata contains a table of paired guide RNAs targeting a paralog pair and their corresponding nucleotide sequences". What's being targeted is a 20bp region of genomic DNA, which is complementary to the gRNA sequence listed

example_pg_metadata <- example_data %>%
  dplyr::select(c("id", "seq_1", "seq_2"))

# sample metadata is the information that describes 
# TODO: ## not sure if the comment above was cut off or not, but more description could help here

example_sample_metadata <- data.frame(
  id = 1:5,
  day = as.factor(c("Day00", "Day05", "Day22", "Day22", "Day22")), 
  rep = as.factor(c("RepA", "RepA", "RepA", "RepB", "RepC"))
)
```

Now let's setup our data using `setup_data()`. Optionally we can provide the metadata in this function as well so that it is stored with the data. 

```{r}
gimap_dataset <- setup_data(counts = example_counts, 
                            pg_metadata = example_pg_metadata,
                            sample_metadata = example_sample_metadata)
```

You'll notice that this set up gives us a list of formatted data. This contains the original counts we gave `setup_data()` function but also normalized counts, and the total counts per sample. 

```{r}
str(gimap_dataset)
```

```{r}
sessionInfo()
```
