---
title: "gimap_annotation_review"
output: html_document
date: "`r Sys.Date()`"
---

```{r}
library(tidyverse)
```

## Load in the GI Mapping annotation table

```{r}
old_annot_df <- readRDS("../extdata/original_d.HeLa_annot")
devtools::load_all()
```

## Get the gimap annotation results

### Get the gimap data setup

```{r}
gimap_dataset <- get_example_data("gimap")
```

### Define the annotation function from the code

```{r}
gimap_dataset <- gimap_dataset %>% 
  gimap_annotate()

new_annot_df <- gimap_dataset$annotation
```

## Compare them

```{r}
nrow(old_annot_df)
nrow(new_annot_df)
```

```{r}
joindf <- dplyr::full_join(old_annot_df, new_annot_df, 
                           by = "pgRNA_id", 
                           suffix = c("_old", "_new"))
```

```{r}
ggplot(joindf, aes(x = log2_cn_gene1, y = gene1_log2_cn)) + 
  geom_point() + 
  ylab("GI_Mapping Output Annotation File: gene1_log2_cn") + 
  xlab("gimap annotation log2_cn_gene1") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1)

```

```{r}
ggplot(joindf, aes(x=log2_cn_gene2, y=gene2_log2_cn)) + 
  geom_point() + 
  ylab("GI_Mapping Output Annotation File: gene2_log2_cn") + 
  xlab("gimap annotation log2_cn_gene2") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1)

```

```{r}
ggplot(joindf, aes(x=log2_tpm_gene2, y=gene2_log2_tpm)) + 
  geom_point() + 
  ylab("GI_Mapping Output Annotation File: gene2_log2_tpm") + 
  xlab("gimap annotation log2_tpm_gene2") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1)

```

```{r}
ggplot(joindf, aes(x=log2_tpm_gene1, y=gene1_log2_tpm)) + 
  geom_point() + 
  ylab("GI_Mapping Output Annotation File: gene1_log2_tpm") + 
  xlab("gimap annotation log2_tpm_gene1") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1)

```

|Which comparison | gimap # NA	| GI Mapping # NA	| # overlap  based on pgRNA ID (using intersect) |
|:--------------:|:------------:|:---------------:|:---------:|
|Gene1 log2 tpm  | `r sum(is.na(new_annot_df$log2_tpm_gene1))
` | `r sum(is.na(old_annot_df$gene1_log2_tpm))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_tpm_gene1))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene1_log2_tpm))]))` |
|Gene 2 log2 tpm | `r sum(is.na(new_annot_df$log2_tpm_gene2))
` | `r sum(is.na(old_annot_df$gene2_log2_tpm))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_tpm_gene2))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene2_log2_tpm))]))` |
|Gene 1 log2 CN	 | `r sum(is.na(new_annot_df$log2_cn_gene1))
` | `r sum(is.na(old_annot_df$gene1_log2_cn))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_cn_gene1))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene1_log2_cn))]))` |
|Gene 2 log2 CN	 | `r sum(is.na(new_annot_df$log2_cn_gene2))
` | `r sum(is.na(old_annot_df$gene2_log2_cn))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_cn_gene2))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene2_log2_cn))]))` |
