---
title: "gimap_annotation_review_afterFilter"
output: html_document
date: "`r Sys.Date()`"
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(ggpubr)
```

## Load in the GI Mapping annotation table

```{r}
old_annot_df <- readRDS("../extdata/original_d.HeLa_annot")
devtools::load_all()
```

## Get the gimap annotation results

### Get the gimap data setup

```{r}
gimap_dataset <- get_example_data("gimap")
```

### Define the annotation function from the code

```{r}
gimap_dataset <- gimap_dataset %>%
  gimap_filter() %>%
  gimap_annotate()

new_annot_df <- gimap_dataset$annotation

old_annot_df <- old_annot_df[which(old_annot_df$pgRNA_id %in% unlist(gimap_dataset$filtered_data$metadata_pg_ids)),]
```

## Compare them

```{r}
nrow(old_annot_df)
nrow(new_annot_df)
```

```{r}
joindf <- dplyr::full_join(old_annot_df, new_annot_df, 
                           by = "pgRNA_id", 
                           suffix = c("_old", "_new"))

joindf_cn <- joindf %>% 
  select(log2_cn_gene1, log2_cn_gene2, gene1_log2_cn, gene2_log2_cn) %>% 
  drop_na()

nrow(joindf_cn)

joindf_tpm <- joindf %>%
  select(log2_tpm_gene1, log2_tpm_gene2, gene1_log2_tpm, gene2_log2_tpm) %>%
  drop_na()

nrow(joindf_tpm)
```

```{r}
annotation_df <- data.frame(label = c(paste0("Number of data points: ", nrow(joindf_cn))),
                            x = 0.25,
                            y = 1.75)
```

```{r}
ggplot(joindf_cn, aes(x = log2_cn_gene1, y = gene1_log2_cn)) + 
  geom_point(alpha=0.1) + 
  ylab("GI_Mapping Output Annotation File: gene1_log2_cn") + 
  xlab("gimap annotation: log2_cn_gene1") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1) +
  geom_text(data = annotation_df, aes(x=x ,y=y, label=label)) +
  stat_regline_equation(label.y=1.55) + stat_cor(label.y = 1.45)

```

```{r}
ggplot(joindf_cn, aes(x=log2_cn_gene2, y=gene2_log2_cn)) + 
  geom_point(alpha=0.1) + 
  ylab("GI_Mapping Output Annotation File: gene2_log2_cn") + 
  xlab("gimap annotation: log2_cn_gene2") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1) +
  geom_text(data = annotation_df, aes(x=x ,y=y, label=label)) +
  stat_regline_equation(label.y=1.55) + stat_cor(label.y = 1.45)
```

```{r}
annotation_df <- data.frame(label = c(paste0("Number of data points: ", nrow(joindf_tpm))),
                            x = 2,
                            y = 10)
```

```{r}
ggplot(joindf_tpm, aes(x=log2_tpm_gene2, y=gene2_log2_tpm)) + 
  geom_point(alpha=0.1) + 
  ylab("GI_Mapping Output Annotation File: gene2_log2_tpm") + 
  xlab("gimap annotation: log2_tpm_gene2") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1) +
  geom_text(data = annotation_df, aes(x=x ,y=y, label=label)) +
  stat_regline_equation(label.y=9) + stat_cor(label.y = 8.25)
```

```{r}
ggplot(joindf_tpm, aes(x=log2_tpm_gene1, y=gene1_log2_tpm)) + 
  geom_point(alpha=0.1) + 
  ylab("GI_Mapping Output Annotation File: gene1_log2_tpm") + 
  xlab("gimap annotation: log2_tpm_gene1") + 
  theme_classic() + 
  geom_abline(intercept = 0, slope = 1) +
  geom_text(data = annotation_df, aes(x=x ,y=y, label=label)) +
  stat_regline_equation(label.y=9) + stat_cor(label.y = 8.25)
```

|Which comparison | gimap # NA	| GI Mapping # NA	| # overlap  based on pgRNA ID (using intersect) |
|:--------------:|:------------:|:---------------:|:---------:|
|Gene1 log2 tpm  | `r sum(is.na(new_annot_df$log2_tpm_gene1))
` | `r sum(is.na(old_annot_df$gene1_log2_tpm))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_tpm_gene1))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene1_log2_tpm))]))` |
|Gene 2 log2 tpm | `r sum(is.na(new_annot_df$log2_tpm_gene2))
` | `r sum(is.na(old_annot_df$gene2_log2_tpm))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_tpm_gene2))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene2_log2_tpm))]))` |
|Gene 1 log2 CN	 | `r sum(is.na(new_annot_df$log2_cn_gene1))
` | `r sum(is.na(old_annot_df$gene1_log2_cn))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_cn_gene1))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene1_log2_cn))]))` |
|Gene 2 log2 CN	 | `r sum(is.na(new_annot_df$log2_cn_gene2))
` | `r sum(is.na(old_annot_df$gene2_log2_cn))` | `r length(intersect(new_annot_df$pgRNA_id[which(is.na(new_annot_df$log2_cn_gene2))]
, old_annot_df$pgRNA_id[which(is.na(old_annot_df$gene2_log2_cn))]))` |
